---
title: "\\textbf{Government - Seattle Resident's Customer Requests}"
author: "Justin Mai"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
  html_document: default
geometry: margin=40pt
fontsize: 11pt
header-includes:
  - \usepackage{titling}
  - \usepackage{titlesec}
  - \titlespacing*{\title}{0pt}{0pt}{0pt}
  - \setlength{\droptitle}{-2em}
  - \setlength{\topskip}{0pt} 
mainfont: "Times New Roman"
---

Goal: Predict the total service requests in the next 3 months for each Service Request Type

After exploring the data, we decided to attempt to predict the total service requests
for the following 3 months. (Multiple Linear Regression intended)

Data Manipulation: After our data exploration and analysis, we decided to remove
zip codes that had null values to work with only service requests that are happening
at a specific location. The removal of these null values removed observations consisting
of online service requests. Even after that, we that there were still 40+ Service Request
Types. To synthesize that down, we broke it into categories

- Vehicle/Parking
- Transportation/Streets
- Environment/Maintenance
- Animal Issues
- Business/Complaints
- City Services/Utilities
- Special Facilities
- Feedback/Other
- Testing

We then removed the Service Request Type and used the Service Request Categories instead
to avoid multicollinearity effects on our model.

```{r}
library(tidyverse)
```

```{r}
df <- read.csv("data/mlr_data.csv")
```

```{r}
df <- df %>%
  mutate(Service_Request_Category = case_when(
    Service_Request_Type %in% c(
      "Abandoned Vehicle", "Abandoned Vehicle/72hr Parking Ordinance", "Parking Enforcement",
      "Public Garage or Parking Lot Complaint", "Towing Complaint - Public Impound", "Towing Impound Complaint"
    ) ~ "Vehicle/Parking",
    
    Service_Request_Type %in% c(
      "Bicycle Facility Maintenance", "Street Sign and Traffic Signal Maintenance", 
      "Street Sign Maintenance", "Streetlight Maintenance", "Traffic Calming", 
      "Traffic Signal Maintenance", "Safe Routes to School"
    ) ~ "Transportation/Streets",
    
    Service_Request_Type %in% c(
      "Clogged Storm Drain", "Damaged Sidewalk", "Dead Animal", "Graffiti",
      "Illegal Dumping / Needles", "Overgrown Vegetation", "Pollution Report Form",
      "Public Litter and Recycling Cans", "Snow and Ice", "Parks and Recreation Maintenance"
    ) ~ "Environment/Maintenance",
    
    Service_Request_Type %in% c(
      "Animal Noise", "Found a Pet", "Lost a Pet", "General Inquiry - Animal Shelter",
      "Nuisance dogs in a park"
    ) ~ "Animal Issues",
    
    Service_Request_Type %in% c(
      "Business Related Complaint", "Business Violation of Public Health Requirements",
      "Taxi, TNC, or Limousine Complaint or Compliment", "Nightlife Noise Complaint"
    ) ~ "Business/Complaints",
    
    Service_Request_Type %in% c(
      "General Inquiry - City Light", "General Inquiry - Customer Service Bureau",
      "General Inquiry - Police Department", "General Inquiry - Public Utilities",
      "General Inquiry - Transportation", "Public Utilities Direct Mail Opt Out",
      "Internet/Cable Issue", "Wireless or Small Cell Issue"
    ) ~ "City Services/Utilities",
    
    Service_Request_Type %in% c(
      "Seattle Center", "Scooter or Bike Share Issue"
    ) ~ "Special Facilities",
    
    Service_Request_Type %in% c(
      "Feedback about the Customer Service Requests Portal", "Feedback about the Find It, Fix It mobile app",
      "Unauthorized Encampment"
    ) ~ "Feedback/Other",
    
    Service_Request_Type %in% c(
      "ZZ-Baseline Testing Only", "ZZDONOTUSE - For Testing Only"
    ) ~ "Testing",
    
    TRUE ~ "Other"  # catch anything unexpected
  ))
```


```{r}

df_grouped <- df %>%
  group_by(Service_Request_Category, Date) %>%
  summarise(Request_Count = sum(Request_Count)) %>%
  ungroup()

ggplot(df_grouped, aes(x = Date, y = Request_Count)) +
  geom_point(alpha = 0.7, color = "#C71585") +
  geom_line(color = "#C71585") +
  facet_wrap(~ Service_Request_Category, scales = "free_y") +
  labs(title = "Service Request Trends by Category",
       x = "Date",   # remove x label
       y = "Request Count") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

```

Our intention was to use Multiple Linear Modeling to attempt to make a prediction however,
the plots above violated too many linear modeling assumptions for it to be accurate.

- Linear Relationship: Within the cluster, there isn't a clear linear relationship between the
number of requests to any category

- Constant Variance: Looking at categories like Transportaion/Streets and Business/Complaints,
there isn't constant variability throughout the plot

The plots with data points for everyday also makes the plots difficult to interpret, leading us
to consolidate down to only using the sum of requests for each Month, Year, and Category, instead
of using Days as well.

```{r}
df1 <- read.csv("data/mlr_data1.csv")

df1 <- df1 %>%
  mutate(Service_Request_Category = case_when(
    Service_Request_Type %in% c(
      "Abandoned Vehicle", "Abandoned Vehicle/72hr Parking Ordinance", "Parking Enforcement",
      "Public Garage or Parking Lot Complaint", "Towing Complaint - Public Impound", "Towing Impound Complaint"
    ) ~ "Vehicle/Parking",
    
    Service_Request_Type %in% c(
      "Bicycle Facility Maintenance", "Street Sign and Traffic Signal Maintenance", 
      "Street Sign Maintenance", "Streetlight Maintenance", "Traffic Calming", 
      "Traffic Signal Maintenance", "Safe Routes to School"
    ) ~ "Transportation/Streets",
    
    Service_Request_Type %in% c(
      "Clogged Storm Drain", "Damaged Sidewalk", "Dead Animal", "Graffiti",
      "Illegal Dumping / Needles", "Overgrown Vegetation", "Pollution Report Form",
      "Public Litter and Recycling Cans", "Snow and Ice", "Parks and Recreation Maintenance"
    ) ~ "Environment/Maintenance",
    
    Service_Request_Type %in% c(
      "Animal Noise", "Found a Pet", "Lost a Pet", "General Inquiry - Animal Shelter",
      "Nuisance dogs in a park"
    ) ~ "Animal Issues",
    
    Service_Request_Type %in% c(
      "Business Related Complaint", "Business Violation of Public Health Requirements",
      "Taxi, TNC, or Limousine Complaint or Compliment", "Nightlife Noise Complaint"
    ) ~ "Business/Complaints",
    
    Service_Request_Type %in% c(
      "General Inquiry - City Light", "General Inquiry - Customer Service Bureau",
      "General Inquiry - Police Department", "General Inquiry - Public Utilities",
      "General Inquiry - Transportation", "Public Utilities Direct Mail Opt Out",
      "Internet/Cable Issue", "Wireless or Small Cell Issue"
    ) ~ "City Services/Utilities",
    
    Service_Request_Type %in% c(
      "Seattle Center", "Scooter or Bike Share Issue"
    ) ~ "Special Facilities",
    
    Service_Request_Type %in% c(
      "Feedback about the Customer Service Requests Portal", "Feedback about the Find It, Fix It mobile app",
      "Unauthorized Encampment"
    ) ~ "Feedback/Other",
    
    Service_Request_Type %in% c(
      "ZZ-Baseline Testing Only", "ZZDONOTUSE - For Testing Only"
    ) ~ "Testing",
    
    TRUE ~ "Other"  # catch anything unexpected
  ))
```

```{r}
df1 <- df1 %>%
  mutate(Date = make_date(Year, Month, 1))

df_grouped1 <- df1 %>%
  group_by(Service_Request_Category, Date) %>%
  summarise(Request_Count = sum(Request_Count)) %>%
  ungroup()

ggplot(df_grouped1, aes(x = Date, y = Request_Count)) +
  geom_point(alpha = 0.7, color = "#C71585") +
  geom_line(color = "#C71585") +
  facet_wrap(~ Service_Request_Category, scales = "free_y") +
  labs(title = "Service Request Trends by Category",
       x = "Date",   # remove x label
       y = "Request Count") +
  theme_minimal() +
  theme(
  )
```

After looking at our plots, we decided to use ARIMA modeling to predict this time series data.
Within most of the categories, there is a clear and relatively stable trend across time which
backs up our use of ARIMA modeling. For categories like Special Facilities, we can see that 
there may be some trends due to seasonal effects, however, since we are only forecasting the
following 3 months, we decided to use a basic ARIMA model, not accounting for seasonal effects.

```{r}
library(forecast)
library(dplyr)
library(lubridate)
library(ggplot2)

# List of categories
categories <- unique(df_grouped1$Service_Request_Category)

# Create an empty list to store results
forecast_list <- list()

# Loop over each category
for (cat in categories) {
  
  # Filter data for one category
  df_cat <- df_grouped1 %>% filter(Service_Request_Category == cat)
  
  # Create time series object
  ts_cat <- ts(df_cat$Request_Count, 
               start = c(year(min(df_cat$Date)), month(min(df_cat$Date))),
               frequency = 12)  # monthly
  
  # Fit ARIMA
  model <- auto.arima(ts_cat)
  
  # Forecast next 3 months
  fc <- forecast(model, h = 3)
  
  # Save the forecast
  forecast_list[[cat]] <- fc
}
```

```{r}
all_forecasts <- bind_rows(
  lapply(names(forecast_list), function(cat) {
    tibble(
      Service_Request_Category = cat,
      Date = seq(max(df_grouped1$Date) %m+% months(1), 
                 by = "month", length.out = 3),
      Predicted_Request_Count = round(as.numeric(forecast_list[[cat]]$mean), 2),
      Lower_80 = round(as.numeric(forecast_list[[cat]]$lower[,1]), 2),
      Upper_80 = round(as.numeric(forecast_list[[cat]]$upper[,1]), 2)
    )
  })
)


ggplot(all_forecasts, aes(x = Date, y = Predicted_Request_Count)) +
  geom_ribbon(aes(ymin = Lower_80, ymax = Upper_80), fill = "#C71585", alpha = 0.3) +  # Confidence interval ribbon
  geom_line(color = "#C71585", size = 1) +  # Forecast line
  facet_wrap(~ Service_Request_Category, scales = "free_y") +  # One plot per service category
  labs(title = "Service Request Forecasts by Category",
       x = "Date",
       y = "Predicted Request Count") +
  theme_minimal()
```

We are making our prediction with an 80% confidence interval. Naturally with ARIMA models,
as we make predictions further from the actual data we have, our confidence interval becomes
wider and wider, this is evident in all of our predictions.

From May to July 2025, we see a lot of variation between predictions of request counts for
each category.

- Animal Issues, Business Complaints, and City Services are all expected to have a constant
number of requests throughout the months. This flat trend is expected in Business/Complaints
and City Services because the prior trend was relatively stable in number of requests per
month. However, there was a clear increasing trend for Animal Issues within recent dates so the flat
forecast could be underestimating the growth in Animal Issues.

- Environment, Feedback, Vehicle/Parking and Special Facilities all forecast an 
increasing trend in the number of request throughout these months. These trends 
could possibly be due to seasonality where as the weather gets warmer, there becomes 
increasingly more complaints about activities and categories concerning the outdoors

- Lastly, Other and Transportation have a decreasing trend throughout the 3 months.
The other category especially has a high confidence interval because we are grouping
many miscellaneous categories together which may not exemplify the same trend. This means
there is expected to be more error in other because of those uncertainties.

Overall, our next steps would be to improve these forecasting predictions by looking deeper
into the model and category separation. We also want to account for seasonality in our
ARIMA model.